---
title: "Silverio_Vasquez_Project1"
author: "Silverio Vasquez"
date: "September 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WTI - Moving Average Crossover Strategy

This R script goes to the EIA website and scrapes the daily WTI prices. A weekly average is calculated, then 13-,26-, and 52-week averages are calculated for a simple moving avreage crossover strategy. 

Then a long/short signal is created if the weekly average is above/below the 52-week average +/- 1 standard deviation.

```{r wti}
# Get packages
#-----------------------------------------------------
# install.packages('rvest')
# install.packages('zoo')
# install.packages('dygraphs')
library(rvest)
library(stringr)
library(zoo)
library(dygraphs)
library(xts)
#-----------------------------------------------------

# Define function
#-----------------------------------------------------
mon_convert <- function(x){
        counter = 1
        limit =13
        mon_az <- c("jan","feb","mar","apr","may","jun","jul","aug",
                    "sep","oct","nov","dec")
        while (counter < limit){
                if (tolower(x) == mon_az[counter]) break
                counter = counter + 1
        }
        if (counter > 12) return(0)
        return(counter)
}
#-----------------------------------------------------
```


Get HTML for EIA website and extract the table headers in the 'B6' CSS class, as well as the WTI spot prices in the 'B3' CSS class.

```{r html}
# Website URL
url <- 'https://www.eia.gov/dnav/pet/hist/LeafHandler.ashx?n=PET&s=RCLC1&f=D'

# Read website HTML
wb <- read_html(url)

# Extract B6 CSS class
wti_html <- html_nodes(wb,'.B6')
wti_text <- html_text(wti_html)

# Extract B3 CSS class
price_html <- html_nodes(wb,'.B3')
price_text <- html_text(price_html)
```


Strip whitespace surrounding text extracted from HTML code. Then convert numbers from characters to numbers, and split year and week into two buckets for 2 columns.

```{r parse}
wti_text <- str_trim(wti_text)
price_text <- as.numeric(price_text)
year_week <- str_split_fixed(wti_text," ",2)
```

Convert extracted data into a dataframe

```{r df}
df <- as.data.frame(matrix(price_text,ncol = 5,byrow = T))
df <- cbind(year_week[,1],year_week[,2],df)
names(df) <- c('year','week','mon','tues','wed','thurs','fri')
head(df)
```

Extract latest date for prices and create a vector of weekly dates to use for plotting, etc.

```{r date}
date <- df$week[dim(df)[1]]
date <- str_split_fixed(date," ",3)[3]
year <- str_trim(df$year[dim(df)[1]])
mon <- str_extract_all(date,"[[:alpha:]]+")
day <- str_extract_all(date,"[0-9]+")

# Use mon_convert function to convert 3 character month into an integer
mon_n <- mon_convert(mon)

date_str <- paste0(year,"/",mon_n,"/",day)

print(paste0('Latest price as of ', date_str))

# Create a vector of weekly dates (each Friday)
idx <- seq(as.Date('1983/4/8'), as.Date(date_str),"week")
head(idx)
```

Calculate moving averages for WTI prices and 52 week standard deviation

```{r sma}
df$wkavg <- apply(df[,3:7],1,mean,na.rm = TRUE)
df$wkavg13 <- rollmean(df$wkavg,13,fill=NA,align='right')
df$wkavg26 <- rollmean(df$wkavg,26,fill=NA,align='right')
df$wkavg52 <- rollmean(df$wkavg,52,fill=NA,align='right')
df$wkstd52 <- rollapply(df$wkavg,52,FUN=sd,fill=NA,align='right')

# Show top of dataframe
head(df)

# Show bottom of dataframe
tail(df)
```

Calculate the log difference of the weekly WTI price aka returns.
Then calculate the position (long or short, +1 or -1) based on whether the weekly average price is above/below the 52-week avg +/- 1 standard deviation. 

```{r signal}
df$wkret <- c(0,diff(log(df$wkavg),lag=1))
df$post <- ifelse(df$wkavg > (df$wkavg52+df$wkstd52),1, ifelse(df$wkavg < (df$wkavg52-df$wkstd52), -1, 0))
```

Make a new dataframe with two columns: weekly returns and long/short signal.
Calculate the strategy's equity curve (the growth of a $1 investment) and plot it.

```{r port}
port <- xts(cbind(df$wkret,df$post),order.by=as.Date(idx))
names(port) <- c('ret','signal')

port$signal <- lag(port$signal,1)
port$signal[is.na(port$signal)] <- 0

port$pnl <- port$ret * port$signal
port$temp <- 0
port$temp <- port$pnl + 1
equity_curve <- cumprod(port$temp)
dygraph(equity_curve,main='Equity Curve') %>% dyRangeSelector()

annRet <- (prod(port$temp)^(1/length(port$temp)))**52*100-100
annStd <- sd(port$temp) * sqrt(52) * 100

print(paste0('The annualized return is ',toString(round(annRet,3)),'%.'))
print(paste0('The annualized standard deviation is ',toString(round(annStd,3)),'%.'))
print(paste0('The Sharpe ratio is ',toString(round(annRet/annStd,3)),'.'))
```

Create a plot with weekly averages, 13-week, 26-week, and 52-week averages.

```{r plot, echo=FALSE}
chart <- xts(cbind(df$wkavg,df$wkavg13,df$wkavg26,df$wkavg52),order.by = as.Date(idx))
names(chart) <- c('WkAvg','13wkAvg','26wkAvg','52wkAvg')
dygraph(chart,main="WTI Moving Averages") %>% dyRangeSelector()
```

